---
title: "Housing Data Exploratory Analysis"
author: "Mark Gingrass"
date: "July 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r housing}
load.libraries <- c('data.table', 'testthat', 'gridExtra', 'corrplot', 'GGally', 'ggplot2', 'e1071', 'dplyr', 'DT')
sapply(load.libraries, require, character = TRUE)
```


## Load train/test Data
```{r, echo = FALSE}
# Bring in some features s characters
train <- fread('input/train.csv',colClasses=c('MiscFeature' = "character", 'PoolQC' = 'character', 'Alley' = 'character'))
test <- fread('input/test.csv' ,colClasses=c('MiscFeature' = "character", 'PoolQC' = 'character', 'Alley' = 'character'))
```

```{r}
cat_var <- names(train)[which(sapply(train, is.character))]
cat_car <- c(cat_var, 'BedroomAbvGr', 'HalfBath', ' KitchenAbvGr','BsmtFullBath', 'BsmtHalfBath', 'MSSubClass')
numeric_var <- names(train)[which(sapply(train, is.numeric))]
```

## Structure of the data
The housing data set has 1460 rows and 81 features with the target feature Sale Price.
Note the structure of the data - treating all factors as characters for now. 
```{r structure}
dim(train)
str(train)

```

## Dimensions
```{r}
dim(train)
DT::datatable(head(train))
```

## Missing Values

```{r}
# Count NAs
na_counts <-lapply(train, function(x) sum(is.na(x)))

na_counts <- melt(na_counts)

names(na_counts) <- c("Number_NAs", "Attribute")
na_counts <- na_counts[which(na_counts$Number_NAs > 0),]

na_counts$Percent_NA <- na_counts$Number_NAs/dim(train)[1]
na_counts <- na_counts[order(-na_counts$Number_NAs),]

DT::datatable((na_counts))
```



PoolQC, MiscFeature, Alley, and Fence have over 50% of data missing in training set. Most likely, I am going to remove those features completley. For the remainder, I am going to impute values. 

```{r missing data_2}
ggplot(na_counts, aes(na_counts$Attribute, na_counts$Number_NAs)) +
  geom_point(shape = 22, colour = "black", fill = "white", size = 2, stroke = 5) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

