---
title: "Explore Housing Data using R"
author: "Mark Gingrass"
date: "August 6, 2018"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE,message=FALSE,error=FALSE,warning=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	error=FALSE,
	message = FALSE,
	warning = FALSE,
	##cache = TRUE,
	tidy = TRUE,
	tidy.opts = list(width.cutoff = 70)
)
```

## Predict SalePrice 

I have learned more about data pipelines from this kaggle kernel linked below. My goal is to improve upon what Aniruddha Chakraborty started.

Some of the exploration and changes I have worked on so far include:

* Using data.table instead of data.frame
* Treating outliers instead of removing them
* Imputing data
* Adding more dialog in some of the methods used for clarity

[Lasso + GBM + XGBOOST - Top 20 % (0.12039) using R](https://www.kaggle.com/aniruddhachakraborty/lasso-gbm-xgboost-top-20-0-12039-using-r)  

# Loading and data pre-processing

## Load data

```{r}
library(data.table)
train <- read.csv("input/train.csv",stringsAsFactors = FALSE)
test <- read.csv("input/test.csv",stringsAsFactors = FALSE)

setDT(train)
setDT(test)

# Size and structure of data sets
dim(train)

# Notice 'test' has one fewer features - SalePrice
dim(test)

str(train)
```

## Data pre-processing

```{r}
## Save Id Columns because they will be dropped form data-set. Id can't be used as a predictor variable.
train_Id <- train$Id
test_Id <- test$Id

## add SalePrice to test data set. 
test$SalePrice <- NA
```

A scatterplot between SalePrice and GrLivArea shows a couple of outliers.

```{r}
library(ggplot2)
qplot(train$GrLivArea, train$SalePrice, main="With Outliers")
```
**Replacing Outliers** - 

Instead of removing outliers, I am replacing them with actual values.
```{r}
# A function to cap outliers
cap_outliers <- function(x){
  qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
  caps <- quantile(x, probs=c(.05, .95), na.rm = T)
  H <- 1.5 * IQR(x, na.rm = T)
  x[x < (qnt[1] - H)] <- caps[1]
  x[x > (qnt[2] + H)] <- caps[2]
  
  return (x)
}

# numeric_df <- dplyr::select_if(train, is.numeric)
train$GrLivArea <- cap_outliers(train$GrLivArea)
train$SalePrice <- cap_outliers(train$SalePrice)
```

```{r}
## Check again after removal.
qplot(train$GrLivArea,train$SalePrice,main="Without Outliers")
```

**Log Transformation of SalePrice Variable** - In order to make the distribution of the target variable normal, we need to transform it by taking log.

```{r}
## histogram of SalePrice - Right skewed
qplot(SalePrice, data = train, bins = 100, main = "Right skewed distribution")

## Log transformation of the target variable
train$SalePrice <- log(train$SalePrice + 1)

## Normal distribution after transformation
qplot(SalePrice,data=train,bins=100,main="Normal distribution after log transformation")
```

**Combine train and test datasets**.

```{r}
## Combine train and test
full_data <- rbind(train,test)
setDT(full_data)

## "Id" unnecessary for prediction
full_data <- full_data[,c("Id") := NULL]
```

# Data Processing and Analysis

## Checking Missing data

Showing the percent of missing data.
```{r}
## Percent of NAs per feature
na_counts <- colSums(is.na(full_data)/dim(full_data)[1] * 100)
na_counts_high <- na_counts[na_counts > 50]
```


## Imputing Missing data

We will be handling each variable separately. 

1. For most of the **categorical features**, NA values will be imputed as **'None'**, because referring to the **data_description.txt** file, **the NA of these variables represent values such as 'No Garage','No Basement', etc.**

2. For most of the **numerical features**, NA values will be replaced by 0, for variables like GarageArea, GarageCars, etc.

3. For some categorical features like Functional and Electrical, the NA values will be replaced by the most frequently occuring value for that variable.

4. Imputing *MSZoning* based on mode of the *Neighborhood* it's in. 


```{r}
## For these features only, fill NA with "None" 
none_vars <- c("Alley","PoolQC","MiscFeature","Fence","FireplaceQu",
               "GarageType","GarageFinish","GarageQual","GarageCond",
               "BsmtQual", "BsmtCond", "BsmtExposure", "BsmtFinType1", 
               "BsmtFinType2", "MasVnrType")

full_data[, (none_vars) := lapply(.SD, function(x) {ifelse(is.na(x), "None", x)}), .SDcols = none_vars]


#Group by neighborhood and fill in missing value by the median LotFrontage of all the neighborhood
temp <- aggregate(LotFrontage~Neighborhood,data = full_data, median)
temp2 <- c()
for(str in full_data$Neighborhood[is.na(full_data$LotFrontage)]) {
  temp2 <- c(temp2,which(temp$Neighborhood == str))
  }

full_data$LotFrontage[is.na(full_data$LotFrontage)]=temp[temp2,2]

## Replacing missing data with 0
zero_vars <- c('GarageYrBlt', 'GarageArea', 'GarageCars','BsmtFinSF1', 'BsmtFinSF2', 'BsmtUnfSF','TotalBsmtSF', 'BsmtFullBath', 'BsmtHalfBath',"MasVnrArea")
full_data[, (zero_vars) := lapply(.SD, function(x) {ifelse(is.na(x), 0, x)}), .SDcols = none_vars]

## Replace missing MSZoning values by mode of neighborhood
Mode <- function(x) {
  ux <- unique(x)
  x <- ux[which.max(tabulate(match(x, ux)))]
}

mode_vars = c("MSZoning")
full_data[, (mode_vars) := lapply(.SD,  function(x) replace(x,
     which(is.na(x)), mode(x))), by = Neighborhood,.SDcols=mode_vars]

## Replace missing Functional values with "Typ"
full_data$Functional[is.na(full_data$Functional)]="Typ"

## Replace missing Electrical values with "SBrkr"
full_data$Electrical[is.na(full_data$Electrical)]="SBrkr"

## Replace missing KitchenQual values by "TA"
full_data$KitchenQual[is.na(full_data$KitchenQual)]="TA"

## Replace missing SaleType values by "WD"
full_data$SaleType[is.na(full_data$SaleType)]="WD"

## Replace missing Exterior1st and Exterior2nd values by "VinylSd"
full_data$Exterior1st[is.na(full_data$Exterior1st)]="VinylSd"
full_data$Exterior2nd[is.na(full_data$Exterior2nd)]="VinylSd"

## All NAs should be gone, except the test portion of SalePrice variable, which we ourselves had initialized to NA earlier.
colSums(is.na(full_data))

full_data$MSSubClass <- as.character(full_data$MSSubClass)
full_data$OverallCond <- as.character(full_data$OverallCond)
full_data$YrSold <- as.character(full_data$YrSold)
full_data$MoSold <- as.character(full_data$MoSold)
```


## Label Encoding some categorical variables that may contain information in their ordering set

**We will also specify the order of the levels (mapping).**

Use this if the categorical variables are in fact known to have ordered relations with the outcome.For example, one type of FireplaceQu might be considered better than another type. 
```{r}
cols = c('FireplaceQu', 'BsmtQual', 'BsmtCond', 'GarageQual', 'GarageCond', 'ExterQual', 'ExterCond','HeatingQC', 'PoolQC', 'KitchenQual', 'BsmtFinType1', 'BsmtFinType2', 'Functional', 'Fence', 'BsmtExposure', 'GarageFinish', 'LandSlope','LotShape', 'PavedDrive', 'Street', 'Alley', 'CentralAir', 'MSSubClass', 'OverallCond', 'YrSold', 'MoSold')

FireplaceQu=c('None','Po','Fa','TA','Gd','Ex')
BsmtQual=c('None','Po','Fa','TA','Gd','Ex')
BsmtCond=c('None','Po','Fa','TA','Gd','Ex')
GarageQual=c('None','Po','Fa','TA','Gd','Ex')
GarageCond=c('None','Po','Fa','TA','Gd','Ex')
ExterQual=c('Po','Fa','TA','Gd','Ex')
ExterCond=c('Po','Fa','TA','Gd','Ex')
HeatingQC=c('Po','Fa','TA','Gd','Ex')
PoolQC=c('None','Fa','TA','Gd','Ex')
KitchenQual=c('Po','Fa','TA','Gd','Ex')
BsmtFinType1=c('None','Unf','LwQ','Rec','BLQ','ALQ','GLQ')
BsmtFinType2=c('None','Unf','LwQ','Rec','BLQ','ALQ','GLQ')
Functional=c('Sal','Sev','Maj2','Maj1','Mod','Min2','Min1','Typ')
Fence=c('None','MnWw','GdWo','MnPrv','GdPrv')
BsmtExposure=c('None','No','Mn','Av','Gd')
GarageFinish=c('None','Unf','RFn','Fin')
LandSlope=c('Sev','Mod','Gtl')
LotShape=c('IR3','IR2','IR1','Reg')
PavedDrive=c('N','P','Y')
Street=c('Pave','Grvl')
Alley=c('None','Pave','Grvl')
MSSubClass=c('20','30','40','45','50','60','70','75','80','85','90','120','150','160','180','190')
OverallCond=NA
MoSold=NA
YrSold=NA
CentralAir=NA

cust_levels <- list(FireplaceQu, BsmtQual, BsmtCond, GarageQual, GarageCond, ExterQual, ExterCond,HeatingQC, PoolQC, KitchenQual, BsmtFinType1, BsmtFinType2, Functional, Fence, BsmtExposure, GarageFinish, LandSlope,LotShape, PavedDrive, Street, Alley, CentralAir, MSSubClass, OverallCond, YrSold, MoSold)
```

I modified a function from [SO](https://stackoverflow.com/questions/38620424/label-encoder-functionality-in-r) to use previously defined levels to encode factors (rather than using sort() to determine levels).
```{r}
# https://stackoverflow.com/questions/38620424/label-encoder-functionality-in-r
# function which returns function which will encode vectors with values  of 'vec' 
lev_index = 1
label_encoder = function(vec){
    levels = cust_levels[[lev_index]]
    lev_index = lev_index + 1
    function(x){
        match(x, levels)
    }
}

full_data[, (cols) := mapply(match, .SD, cust_levels, SIMPLIFY = FALSE), .SDcols = cols]
```
